Drought-sensitive trees survive on deep-water access in a tropical forest community

### 2020-08-09 manuscript.rmd
model comparison moved back to methods

---
title: "Drought-sensitive trees survive on deep-water access in a tropical forest community"
author: ''
date: ''
output:
  bookdown::word_document2:
        reference_docx: manuscriptstyle.docx
        fig_caption: yes
  bookdown::html_document2:
    number_sections: no
    theme: flatly
    fig_caption: yes
  bookdown::pdf_document2: default
bibliography: references.bib
csl: new-phytologist.csl
---

@@

@@

\
In comparison to the measured isotopic signal, our results showed that the model with VPD and without Leaf Cover (Eq. \@ref(eq:M2)) had highest goodness-of fit (*R*^2^ = 0.51; *p* = 0.072), followed by followed by model without both VPD and Leaf Cover (Eq. \@ref(eq:M1); Fig. S\@ref(fig:other-erd-mods)). Models with Leaf Cover had only three species with sufficient species-level goodness-of-fit (Eqs. \@ref(eq:M3)-\@ref(eq:M4)) versus seven for models without (Eqs. \@ref(eq:M1)-\@ref(eq:M2)). For rest of the analyses we therefore used species ERD estimates from the model with VPD and without Leaf Cover (Eq. \@ref(eq:M2); Fig. \@ref(fig:erd-iso)).

@@

\
In this forest modelled effective rooting depths (ERD) for the `r length(erd.sp)` large, canopy tree species varied from 0.5 m to 13 m (Fig. \@ref(fig:sp-erd)). For the subset of species for which hydraulic properties were available, species with deeper ERD showed greater maximum stem hydraulic conductance ($K_{max,~stem}$; *R*^2^ = 0.39, *P* = 0.098, *n* = 8; Fig. \@ref(fig:erd-hyd)A), less negative leaf ($\Psi_{tlp}$); *R*^2^ = 0.6, p = 0.024, *n* = 8; Fig. \@ref(fig:erd-hyd)B), greater vulnerability to embolism from cavitation with less negative ($\Psi_{88,~stem}$) (*R*^2^ = 0.72, p = 0.007, *n* = 8; Fig. \@ref(fig:erd-hyd)C), and narrower hydraulic safety margins ($\Psi_{min} - \Psi_{88,~stem}$; *R*^2^ = 0.74, p = 0.013, *n* = 8; Fig. \@ref(fig:erd-hyd)D).

```{r sp-erd, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "60%", fig.pos = 'H', results = "asis", fig.cap= paste0("\\label{lab-sp-erd}Modelled Effective Rooting Depth (ERD; mean $\\pm$ 1SE in m) for ", length(erd.sp)," large, canopy species of Barro Colorado Island. See Table  \\@ref(tab:spcodes) for scientific names corresponding to species codes.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/ERD_by_sp_large_canopy.jpeg"))
```

@@

```{r iso-data, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "40%", fig.pos = 'H', results = "asis", fig.cap = "\\label{lab-iso-data}Stable isotopic concetration ($\\delta^2H$) of soil water by depth measured in March 1997. Data from @Meinzer:1999."}
knitr::include_graphics(file.path("figures/UDI_confidence/Meinzer_etal_1999_Depth_vs_soil_deltaD.jpeg"))
```

@@

```{r dbh-model, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', dpi = 96, out.width = "60%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-dbh-model} A scatterplot of tree growth rates (circles) from all census intervals versus dbh at the beginning of the cenus interval is overlaid with a dbh model of growth (spline). Growth rates are color coded by census interval. The model equation is shown in inset along with goodness-of-fit(*R*^2^)")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/dbh/growthrate_dbh_predicted_growth_with_spline_degree_5_dbh_cutoff_100_model.jpeg"))
```

>>>>

---
title: "Drought-sensitive trees survive on deep-water access in a tropical forest community"
author: ''
date: ''
output:
  bookdown::html_document2:
    number_sections: no
    theme: flatly
    fig_caption: yes
  bookdown::word_document2:
        reference_docx: manuscriptstyle.docx
        fig_caption: yes
  bookdown::pdf_document2: default
bibliography: references.bib
csl: new-phytologist.csl
---

@@

The model with VPD and without Leaf Cover had highest goodness-of fit (*R*^2^ = 0.51; *p* = 0.072), followed by model without both VPD and Leaf Cover (Eq. \@ref(eq:M1); Fig. S\@ref(fig:other-erd-mods)). Models with Leaf Cover had only three species with sufficient species-level goodness-of-fit (Eqs. \@ref(eq:M3)-\@ref(eq:M4)) versus seven for models without (Eqs. \@ref(eq:M1)-\@ref(eq:M2)). For rest of the analyses we therefore used species ERD estimates from the model with VPD and without Leaf Cover (Eq. \@ref(eq:M2)).

@@

\
In comparison to the measured isotopic signal, our results showed that the ERD model with VPD and without Leaf Cover (Eq. \@ref(eq:M2)) had highest goodness-of fit (*R*^2^ = 0.51; *p* = 0.072) and was thus the chosen model of ERD for rest of the analyses. Notably, two species, *Luehea* *seemannii* and *Trichilia tuberculata*, that @Meinzer:1999 found to have xylem sap $\delta^2H$ values between soil water and groundwater, suggesting that these species sourced most of the water from depths > 1 m, were also identified by our model with ERD > 1 m (2.9 m for both species). (Fig. \@ref(fig:erd-iso)).

@@

\
In this forest modelled effective rooting depths (ERD) for the `r length(erd.sp)` large, canopy tree species varied from `r min(df.erd.to.plot$depth)` m to `r max(df.erd.to.plot$depth)` m (Fig. \@ref(fig:sp-erd)). For the subset of species for which hydraulic properties were available, species with deeper ERD showed greater maximum stem hydraulic conductance ($K_{max,~stem}$; *R*^2^ = 0.39, *P* = 0.098, *n* = 8; Fig. \@ref(fig:erd-hyd)A), less negative leaf ($\Psi_{tlp}$); *R*^2^ = 0.6, p = 0.024, *n* = 8; Fig. \@ref(fig:erd-hyd)B), greater vulnerability to embolism from cavitation with less negative ($\Psi_{88,~stem}$) (*R*^2^ = 0.72, p = 0.007, *n* = 8; Fig. \@ref(fig:erd-hyd)C), and narrower hydraulic safety margins ($\Psi_{min} - \Psi_{88,~stem}$; *R*^2^ = 0.74, p = 0.013, *n* = 8; Fig. \@ref(fig:erd-hyd)D).

```{r sp-erd, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "60%", fig.pos = 'H', results = "asis", fig.cap= paste0("\\label{lab-sp-erd}Modelled Effective Rooting Depth (ERD; mean $\\pm$ 1SE in m) for ", length(erd.sp)," large, canopy species of Barro Colorado Island. See Table  \\@ref(tab:spcodes) for complete scientific names.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/ERD_by_sp_large_canopy.jpeg"))
```

@@

```{r iso-data, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "40%", fig.pos = 'H', results = "asis", fig.cap = "\\label{lab-iso-data}Stable isotopic ratio ($\\delta^2H$) of soil water by depth measured at Barro Colorado Island in March 1997. Data from @Meinzer:1999."}
knitr::include_graphics(file.path("figures/UDI_confidence/Meinzer_etal_1999_Depth_vs_soil_deltaD.jpeg"))
```

@@

```{r dbh-model, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', dpi = 96, out.width = "60%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-dbh-model} A scatterplot of tree growth rates (circles) from all census intervals versus dbh at the beginning of the cenus interval is overlaid with a dbh model of growth (spline). Growth rates are color coded by census interval. The model equation is shown in inset along with goodness-of-fit(*R*^2^)")}
knitr::include_graphics(file.path("figures/dbh/growthrate_dbh_predicted_growth_with_spline_degree_5_dbh_cutoff_100_model.jpeg"))
```
